---
title: "Exercise"
output: 
 html_document:
    toc: true
    #number_sections: true
    toc_float: true
    collapsed: false
---
Starting from the raw fishing data of 5 vessels operating in GSA 23 apply the R4ais workflow to obtain fishing segments, and then analyse the data to obtain a map of the fishing effort. The exercise is divided into 4 questions which helps you to get the desired result. All the code you need is part of the script seen today.


# Before to start

You need to check your working directory (WD), which have to be "dwrs_workshop_2022". The `getwd('..')` command shows you where the WD is. The `setwd('..')` command set the WD to a backward place. 

```{r cars, echo=T, message=F, warning=F}
getwd() # The WD have to be "dwrs_workshop_2022"
setwd('..') # Push one to set the WD to a backward place
getwd() # check the WD: it have to be "practical_session_2. If is not yet right, push again on line 6 and check"
```
```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = 'C:/Users/e.armelloni/OneDrive/Lavoro/AIS/Lavori/Lavori_2022/GFCM_workshop/dwrs_workshop_2022')
library(kableExtra)

```

### Load sample data

Data needed for the exercise are the sample fishing points

```{r, echo=T}
sample_data=readRDS("practical_session_2/data/exercise_data.rData")
```
```{r, echo=F}
head((sample_data))%>%
  kbl() %>%
  kable_paper("hover", full_width = F)
```


```{r, echo=F, message=F,warning=F}
library(leaflet)
library(RColorBrewer)
library(sf)
pal_fun <- colorQuantile("YlOrRd", NULL, n = 4)
st_as_sf(sample_data, coords = c('longitude' ,'latitude' )) %>%
  leaflet( width = 900) %>%addCircles(
   # remove polygon borders
    
   opacity = 0.5,
   weight =3,
   color='red')%>%
  addProviderTiles(providers$Esri.OceanBasemap) 
  #addTiles()


```

### Load R4ais functions

In addition you need to load the functions contained in the R4ais release. You need just to run the following lines

```{r, echo=T, message=F,warning=F}
base_dir=getwd()
setwd(file.path(base_dir, 'practical_session_1'))
dirmaps="AIS_data_processing/maps" # path of the maps directory
file_parameters="AIS_data_processing/data/parameters.csv" # import parameter table
file_centroids="AIS_data_processing/data/centroids.csv" # import centroids 
wgs="+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0" # Insert your coordinates system
install.missing.packages=T # set to TRUE if want to allow automatic installation of missing packages.
source("AIS_data_processing/R/global_functions.R") # load all internal function required
ports<-read_sf(file.path(dirmaps, "med_harb_gsa")) # import list of ports
port_buf<-st_buffer(ports, 0.001) # create a buffer
st_crs(port_buf)=wgs # set crs
coastal_ban_zone=read_sf(file.path(dirmaps, "coastal_ban_zone")) # import managment depth layer
st_crs(coastal_ban_zone)=wgs # set crs
grid<-read_sf(file.path(dirmaps, "grid01degrees")) # import a grid
grid$grid_id=seq(1:nrow(grid)) # create cell grid id
st_crs(grid)=wgs # set crs
centroids=inport_parameters(file.path(file_parameters), file.path(file_centroids))[[2]]
pars=inport_parameters(file.path(file_parameters), file.path(file_centroids))[[1]]
setwd(base_dir)
```

<br>

# Questions

### Q1
Obtain the fishing segments from one single vessel (data provided): you need to identify the right function from the workflow.
Tip: the output of the function is a list, so you need to process the output as x=plyr::ldply(x)

As input data you can use sample_dat_Q1, produced as:

```{r, echo=T}
sample_dat_Q1=sample_data[sample_data$MMSI==unique(sample_data$MMSI)[1],]
```


### Q2
Extract fishing segments from all the vessels in the sample data at once: you need to identify the right function from the workflow. The resulting object must be called `all_fishing_tracks`
Tip: after having applied the right function, you need to pass the output to the following lines 

```{r, echo=F}
all_fishing_tracks=list(a=list())
```

```{r, echo=T}
# post process 1
for(i in 1:length(all_fishing_tracks)){
all_fishing_tracks[[i]]=plyr::ldply(all_fishing_tracks[[i]])
}
# post process 2
all_fishing_tracks=plyr::ldply(all_fishing_tracks)

```



### Q3
Intersect the obtained data with the grid of GSA 23. Tip: the piece of code you need is a for loop.

Before to start, run the following lines to import the correct function and parameters (attention: this is an extra tip).
```{r, echo=T}
source("practical_session_2/R/supporting_code/global_functions_light.R")
grid=read_sf("practical_session_2/maps/grid")
activate_filter='Y'
ref_years=2015:2018
```


### Q4
Create a map of the fishing effort.
Tip: the following lines import all the shapefiles and the plot specifications.

```{r, echo=T, message=F,warning=F}
grid_whole <- read_sf("practical_session_2/maps/grid")

land=read_sf("practical_session_2/maps/land")%>% 
  st_crop(., xmin=-10, xmax=43, ymin=20, ymax=48) #

gsa<-read_sf("practical_session_2/maps/gsa")%>%
  dplyr::filter(SMU_CODE == '23' )

dep = read_sf("practical_session_2/maps/depth_contours")%>%
  st_cast(., to = "POLYGON")%>%
  st_set_crs(st_crs(gsa))%>%
  st_buffer(0)%>%
  st_crop(gsa)

lims=c(st_bbox(gsa[c(1,3)], st_bbox(gsa[c(2,4)])))

#### Specifications for the plot ####
# Plot size
text_size=8
plot_height=25
plot_width=18
library(ggspatial)
library(ggnewscale)
source("practical_session_2/R/supporting_code/plot_specifications.R")
```




